---
title: "lineagespot User Guide"
author:
- name: Nikolaos Pechlivanis
  affiliation: Institute of Applied Biosciences, 
    Centre for Research and Technology Hellas, Thessaloniki, GR
  email: nikosp41@certh.gr
- name: Maria Tsagiopoulou
  affiliation: Institute of Applied Biosciences, 
    Centre for Research and Technology Hellas, Thessaloniki, GR
- name: Maria Christina Maniou
  affiliation: Institute of Applied Biosciences, 
    Centre for Research and Technology Hellas, Thessaloniki, GR
- name: Anastasis Togkousidis
  affiliation: Institute of Applied Biosciences, 
    Centre for Research and Technology Hellas, Thessaloniki, GR
- name: Evangelia Mouchtaropoulou
  affiliation: Institute of Applied Biosciences, 
    Centre for Research and Technology Hellas, Thessaloniki, GR
- name: Taxiarchis Chassalevris
  affiliation: School of Veterinary Medicine, 
    Aristotle University of Thessaloniki, GR
- name: Serafeim Chaintoutis
  affiliation: School of Veterinary Medicine, 
    Aristotle University of Thessaloniki, GR
- name: Chrysostomos Dovas
  affiliation: School of Veterinary Medicine, 
    Aristotle University of Thessaloniki, GR
- name: Maria Petala
  affiliation: Dept. of Civil Engineering, 
    Aristotle University of Thessaloniki, GR
- name: Margaritis Kostoglou
  affiliation: Dept. of Chemistry, 
    Aristotle University of Thessaloniki, GR
- name: Thodoris Karapantsios
  affiliation: Dept. of Chemistry, 
    Aristotle University of Thessaloniki, GR
- name: Stamatia Laidou
  affiliation: Institute of Applied Biosciences, 
    Centre for Research and Technology Hellas, Thessaloniki, GR
- name: Elisavet Vlachonikola
  affiliation: Institute of Applied Biosciences, 
    Centre for Research and Technology Hellas, Thessaloniki, GR
- name: Anastasia Chatzidimitriou
  affiliation: Institute of Applied Biosciences, 
    Centre for Research and Technology Hellas, Thessaloniki, GR
- name: Agis Papadopoulos
  affiliation: EYATH S.A., 
    Thessaloniki Water Supply and Sewerage Company S.A., Thessaloniki, GR
- name: Nikolaos Papaioannou
  affiliation: School of Veterinary Medicine, 
    Aristotle University of Thessaloniki, GR
- name: Anagnostis Argiriou
  affiliation: Institute of Applied Biosciences, 
    Centre for Research and Technology Hellas, Thessaloniki, GR
- name: Fotis E. Psomopoulos
  affiliation: Institute of Applied Biosciences, 
    Centre for Research and Technology Hellas, Thessaloniki, GR

date: "`r BiocStyle::doc_date()`"
package: "`r BiocStyle::pkg_ver('lineagespot')`"
output: 
  BiocStyle::html_document:
    toc_float: true
vignette: >
  %\VignetteIndexEntry{lineagespot User Guide}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
```{r vignetteSetup, echo=FALSE, message=FALSE, warning = FALSE}
## For links
library("BiocStyle")
## Track time spent on making the vignette
startTime <- Sys.time()
## Bib setup
library("RefManageR")
## Write bibliography information
bib <- c(
    R = citation(),
    BiocStyle = citation("BiocStyle")[1],
    vcfR = citation("vcfR")[1],
    data.table = citation("data.table")[1],
    stringr = citation("stringr")[1]
)
```


<style>
<!-- h1, h2, h3, h4 { -->
<!--   color:#17247a; -->
<!-- } -->

strong {
    color:#eb6b1c;
}

</style>

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
    comment = "#>",
    error = FALSE,
    warning = FALSE,
    message = FALSE,
    crop = NULL
)
```
# Introduction

Lineagespot is a framework written in [R](https://www.r-project.org/), which aims 
to identify SARS-CoV-2 related mutations based on a single (or a list) of 
variant(s) file(s) (i.e., [variant calling format](https://gatk.broadinstitute.org/hc/en-us/articles/360035531692-VCF-Variant-Call-Format)). 


## Installation

`lineagespot` is distributed as a [Bioconductor](https://www.bioconductor.org/) 
package and requires `R` (version "4.1"), which can be installed on any 
operating system from [CRAN](https://cran.r-project.org/), and 
Bioconductor (version "3.14").

To install `lineagespot` package enter the following commands in 
your `R` session:

```{r eval = FALSE}
if (!requireNamespace("BiocManager", quietly = TRUE)) {
    install.packages("BiocManager")
}
BiocManager::install("lineagespot")
## Check that you have a valid Bioconductor installation
BiocManager::valid()
```

## Launching the app

Once `lineagespot` is successfully installed, it can be loaded as follow:

```{r setup}
library(lineagespot)
```


# Session info {.unnumbered}

Here is the output of `sessionInfo()` on the system on which this document was
compiled running pandoc ``r rmarkdown::pandoc_version()``:

```{r sessionInfo, echo=FALSE}
sessionInfo()
```

